### 2023-12-27 thesis/chapters/01-introduction.tex
Move machinery and sensors to implementation and resume background

\chapter{Introduction}
Manufacturing experiences a shift in traditional asset operational status evaluation and utilization. The rise of Industry 4.0 means greater automation and robotization of the production halls to achieve optimal usage of available resources. The secondary aspect in the enterprises' endeavor, however not less important, is to keep track of the equipment wear and tear. The corrective action be it repair or replacement should be done on time in response to the key indicators.

@@

Vibrations are the most non-intrusive way to sense and record eventually fatal deficiencies right at the onset. The experts use it to distinguish faulty states and to identify the malfunction's root causes. In critical circumstances such as is the case of the large turbines in the power plants, the precautions leading to regular machinery check-ups are already in place. The monitoring solution has to be sufficiently independent, reliable, and accurate to reach wider acceptance and spread.

@@

Chapter \ref{chapter:design} on solution design starts by posing research questions and establishing goals. Then, the MaFaulDa dataset statistical properties and labeling procedure are described on the original samples as well as on extracted attributes. Data rates with feature selection are derived based on the MaFaulDa. In the section about classification models, we compare the performance for various feature sets in the k-nearest neighbors algorithm. Finally, in preparation for the evaluation part, designated machines in the industrial environment are presented alongside the sensor device suitable for their measurement.

>>>>

\chapter{Introduction}
The industry experiences a shift in traditional asset operational status evaluation and utilization. The rise of Industry 4.0 means greater automation and robotization of the production halls to achieve optimal usage of available resources. The secondary aspect in the enterprises' endeavor, however not less important, is to keep track of the equipment wear and tear. The corrective action be it repair or replacement should be done on time in response to the key indicators.

@@

Vibrations are non-intrusive way to sense and record eventually fatal deficiencies right at the onset. The experts use it to distinguish faulty states and to identify the malfunction's root causes. In critical circumstances such as is the case of the large turbines in the power plants, the precautions leading to regular machinery check-ups are already in place. The monitoring solution has to be sufficiently independent, reliable, and accurate to reach wider acceptance and spread.

@@

Chapter \ref{chapter:design} on solution design starts by posing research questions and establishing goals. Then, the MaFaulDa dataset statistical properties and labeling procedure are described on the original samples as well as on extracted attributes. Data rates with feature selection are derived based on the MaFaulDa. In the section about classification models, we compare the performance for various feature sets in the k-nearest neighbors algorithm. Finally, in preparation for the evaluation part, the sensor device is propresed which is suitable for measurements of designated machines in the industrial environment.

### 2023-12-27 thesis/chapters/03-design.tex
Introduction
Move machinery and sensors to implementation and resume background

\chapter{Design} \label{section:design}
The design phase sets out to elaborate on the scenarios for experiments with IoT device. In these experiments, we compare and combine approaches from industry standards and papers presented in the analysis section. Conclusions are taken into consideration in establishing the plan of measurements and in the construction of the sensor unit.

@@

\section{Vibration measurement plan}
%TODO
Identify machines, measurement positions, vibration sensor requirements and parameters with schedule for long-term measurement. It has to take into account with machinery operations.

\subsection{Machinery}
We have three kinds of rotating machines at disposal for measurements which are standing fan, scroll compressor, and water pump:
\begin{itemize}
\itemsep0pt
\item \textbf{Standing fan}: model \emph{Kalorik TKG VT 1037 C} (Fig.~\ref{fig:machine:fan}) 1 piece with 1 measurement axial position on the plastic casing back of the electromotor  as testing environment for firmware. It has 45W of power (ISO 20186 class \rom{1}) and 45 cm fan diameter consiting of 3 propelers. 3 speeds are approximated with tolerance of $\pm$ 0.61 Hz (37 rpm) by measuring fundamental frequency to speed 1 is 18.5 Hz (1100 rpm), speed 2 is 20.4 Hz (1200 rpm), speed 3 is 22 Hz (1300 rpm). Acceleration observed is $\pm$ 5 $\mathrm{m/s}^2$ in radial and tangential directions, and $\pm$ 1 $\mathrm{m/s}^2$ in axial direction.

\item \textbf{Scroll compressor}: model \emph{Copeland ZR16M3E-TWD-561} (Fig.~\ref{fig:machine:compressor}) 2 pieces in two independent air condition units Vertiv for data center. Compressor has 9.7 kW of power (ISO20186 class \rom{1}). It rotates with 2900 rpm (48.3 Hz). Faulty parts can result in unbalance of the rotor, fault of the scroll, mechanical assembly loose, or losse bearing. It has two possible measuremnt locations from sides on bearing just above the base and above the scroll. Accelerations picked up by sensor are $\pm$ 3 $\mathrm{m/s}^2$ tops.

\item \textbf{Water pump}: 3 municipal single-stage axially split volute casing pump and attached electric induction motor with squirrel cage rotor. Newer primary pump with vendor cloud monitoring solution is \emph{KSB Omega 300-560 B GB G F} (Fig.~\ref{fig:machine:pump-ksb}) commisioned in 2019, with rotational speed around 1500 rpm (precisly 1493 rpm or 24.9 Hz). Motor has 400 kW of power (ISO20186 class \rom{2}). Outer bearing on pump experiences acceleration of  Secondary pumps are  two pieces produced by Sigma (Fig.~\ref{fig:machine:pump-sigma}) in 1986 with motor powers of $\pm$ 5 $\mathrm{m/s}^2$ ,  . Each pump and motor has 2 bearing therefore 4 measuement positions in total.
\end{itemize}

\begin{figure}[ht]
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
    		\centering
        \includegraphics[width=0.4\textwidth]{assets/design/machine-fan.jpg}
        \caption{Stand fan Kalorik}
        \label{fig:machine:fan}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
    		\centering
        \includegraphics[width=0.4\textwidth]{assets/design/machine-compressor.jpg}
        \caption{Scroll compressor Copeland}
        \label{fig:machine:compressor}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
    		\centering
        \includegraphics[width=0.8\textwidth]{assets/design/machine-ksb-pump.jpg}
        \caption{Water pump KSB Omega}
        \label{fig:machine:pump-ksb}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
    		\centering
        \includegraphics[width=0.8\textwidth]{assets/design/machine-sigma-pump.jpg}
        \caption{Water pump Sigma}
        \label{fig:machine:pump-sigma}
    \end{subfigure}
    \caption{Machines dedicated for vibration measurements}
\end{figure}

%TODO bi-weekly measurements. Describe preliminary measurements and orbital

\begin{figure}[ht]
    \centering
    \begin{subfigure}[b]{0.44\textwidth}
        \includegraphics[width=\textwidth]{assets/design/EDA-custom-dataset-temporal.png}
        \caption{Temporal domain waveforms in all spatial directions with 300 ms duration}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.55\textwidth}
        \includegraphics[width=\textwidth]{assets/design/EDA-custom-dataset-spectral-X-axis.png}
        \caption{Spectral domain in x accelerometer axis. FFT with Hann window of length $2^{10}$ ($\approx 409$ ms)}
    \end{subfigure} 
    \caption{Signals at 5 s timestamp}
\end{figure}

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{assets/design/compressor-orbitals-1x.png}
	\caption{Orbital analysis of compressor's 1x harmonic}
\end{figure}

\subsection{Sensors}
In prelimary exploratory vibration measurements of machines we used ADXL335 accelerometer connected to BeagleBone Black microcontroller. Hardware characteristics are depicted in Table~\ref{tab:design:hw-sensors}. The bandwidth, sensitivity and  sampling frequency in the first attempt solution we find to be not adequate (satisfactory).

\begin{table}[ht]
\renewcommand{\arraystretch}{1.2}
\centering
\begin{tabular}{|l|l|l|}
\hline
\textbf{Accelerometer}                           & \textbf{ADXL335} & \textbf{IIS3DWB}   \\ \hline
\textbf{Vendor}                                  & Analog Devices   & STMicroelectronics \\ \hline
\textbf{Bus}                                     & Analog           & SPI                \\ \hline
\textbf{Axis}                                    & 3                & 1 or 3             \\ \hline
\textbf{Range} (g)                               & $\pm$ 3          & $\pm$ 2 to 16      \\ \hline
\textbf{Bandwidth} (kHz)                         & 0.55             & 5 - 6.3            \\ \hline
\textbf{Sensitivity}                             & 300 mV/g         & 0.061 mg/LSB       \\ \hline
\textbf{Noise density} ($\mu g / \sqrt{\mathrm{Hz}}$ rms) & 150 - 300        & 75                 \\ \hline
\textbf{Microcontroller}                         & Beaglebone Black & ESP32-PoE-ISO      \\ \hline
\textbf{CPU SoC}                                 & TI Sitara AM3358 & ESP32-WROOM-32     \\ \hline
\textbf{Output data rate} (kHz)                  & 2.5              & 26.7               \\ \hline
\textbf{ADC resolution} (bit)                    & 12               & 16                 \\ \hline
\textbf{FIFO}                                    & -                & 3 kB (512 samples) \\ \hline
\end{tabular}
\caption{Accelerometers}
\label{tab:design:hw-sensors}
\end{table}

The ST IIS3DWB is appropriately low priced accelerometer approching industrial standards for vibration monitoring. We need high speed interface to store all recorded samples to be used in later analysis and optimization. Therefore as a microcontroller development kit we choose ESP32-PoE-ISO that has SD card slot connected via SD/MMC bus. SPI bus can be connected to any GPIO pin and is routed via GPIO matrix at lower but enough speed, sensor can handle SPI speed of 10 MHz at maximum. The block diagram of the designed hardware device is in Figure~\ref{fig:design:block-diagram-hw}.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.9\textwidth]{assets/design/hw-block-schematic.png}
	\caption{Sensor unit hardware block diagram}
	\label{fig:design:block-diagram-hw}
\end{figure}

The drivers for neccessary low level interfaces and FatFS filesystem are already implemented in ESP-IDF software development kit\footnote{ESP-IDF SDK: \url{https://docs.espressif.com/projects/esp-idf/en/latest/esp32/api-reference/index.html}}. The acceleromter driver is also made available by vendor and can be integrated with just two function calls to bus read and write operations\footnote{IIS3DWB driver: \url{https://github.com/STMicroelectronics/iis3dwb-pid}}.

>>>>

\chapter{Design} \label{chapter:design}
The design phase sets out to elaborate on the scenarios for experiments with IoT device. In these experiments, we compare and combine approaches from industry standards and papers presented in the analysis section. Conclusions are taken into consideration in establishing the plan of measurements and in the construction of the sensor unit.

@@

### 2023-12-27 thesis/chapters/04-implementation.tex
Move machinery and sensors to implementation and resume background

\chapter{Implementation} \label{section:implementation}

>>>>

\chapter{Implementation} \label{section:implementation} 
The methodology applied so far for the dataset recorded in a laboratory is to be tested on vibration signals from the industrial environment. Established guidelines for introducing a monitoring system have been described in the section on technical standards (Section~\ref{section:technical-standards}).

We employ a slightly customized validation procedure that involves the selection of machines to be monitored, identifying positions for measurements according to technical specifications, taking preliminary readings, and developing a device capable of capturing the usual failure modes. The collection of the novel dataset is accompanied by an agreed upon schedule.

\section{Monitored machinery}
Three kinds of rotating machines are at disposal for vibration measurements. Those are a standing fan, scroll compressor, and water pump.

\begin{figure}[ht]
    \centering
    \begin{subfigure}[b]{0.17\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{assets/design/machine-fan.jpg}
        \caption{\footnotesize Fan}
        \label{fig:machine:fan}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.17\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{assets/design/machine-compressor.jpg}
        \caption{\footnotesize Compressor}
        \label{fig:machine:compressor}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.31\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{assets/design/machine-ksb-pump.jpg}
        \caption{\footnotesize Water pump KSB}
        \label{fig:machine:pump-ksb}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.31\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{assets/design/machine-sigma-pump.jpg}
        \caption{\footnotesize Water pump Sigma}
        \label{fig:machine:pump-sigma}
    \end{subfigure}
    \caption{Machines dedicated for vibration measurements}
\end{figure}

\textbf{Standing fan} is model \emph{Kalorik~TKG~VT1037} (Fig.~\ref{fig:machine:fan}) and one unit is available to us. It serves as a test bed during the sensor unit development. The sensor is placed on the plastic casing at the back of the drive motor. The fan has a 45~cm diameter with 3~propellers and a power of 45~W (class~\rom{1}). It has a switch for 3 rotational speeds, which are approximately 18.5~Hz (1100~rpm), 20.4~Hz (1200~rpm), and 22~Hz (1300~rpm). Acceleration observed is $\pm 5\;\mathrm{m/s}^2$ in radial and tangential directions, and $\pm 1\;\mathrm{m/s}^2$ in axial direction.

\textbf{Scroll compressor} is model \emph{Copeland~ZR16} (Fig.~\ref{fig:machine:compressor}) and altogether two units are available in two independent air conditioning units for the data center. The compressor has 9.7~kW of power (class~\rom{1}) and rotates at 2900 rpm (48.3 Hz). Two possible measurement locations are located on the sides on top of the bearings, just above the base and above the scroll. Accelerations picked up by the sensor are $\pm 3\;\mathrm{m/s}^2$ tops.

\textbf{Water pumps} are available as 3 units in municipal drinking water pumping station. The apparatus consists of a single-stage axially split volute casing pump and an attached electric induction motor with a squirrel cage rotor.

The newer primary pump with cloud monitoring solution is one unit of \emph{KSB Omega~300-560} (Fig.~\ref{fig:machine:pump-ksb}). The pump was commissioned in 2019 and it rotates at a speed of circa 1500 rpm (precisely 1493 rpm or 24.9 Hz). The electric motor in the vicinity has 400~kW of power (class~\rom{2}). Outer bearing on pump vibrates with acceleration of $\pm 5 \;\mathrm{m/s}^2$.

Secondary pumps are two units produced by \emph{Sigma} (Fig.~\ref{fig:machine:pump-sigma}) in 1986 with motor powers of 250~kW and 100~kW. Each pump and motor has 2 bearings therefore there are 4 measurement positions in total.

\textbf{Complete schedule} for planned vibration data gathering on the designated machines and measurement procedure is described in Appendix \ref{appendix:data-collection}. In addition, we might be able to export the entire historical logs of the water pump from \textbf{KSB cloud} to enlarge the dataset.

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{assets/design/ksb-cloud.jpg}
     \caption{KSB cloud pump monitoring after ramp-up with displayed warning and alert levels}
\end{figure}

\section{Sensor hardware and drivers}
In preliminary exploratory vibration measurements of machines, the ADXL335 accelerometer is connected to BeagleBone Black microcontroller. The bandwidth, sensitivity, and sampling frequency of the makeshift solution are not appear to be adequate (Table~\ref{tab:design:hw-sensors}).

\begin{table}[ht]
\renewcommand{\arraystretch}{1.2}
\centering
\begin{tabular}{|l|l|l|}
\hline
\textbf{Accelerometer}                           & \textbf{ADXL335} & \textbf{IIS3DWB}   \\ \hline
\textbf{Vendor}                                  & Analog Devices   & STMicroelectronics \\ \hline
\textbf{Bus}                                     & Analog           & SPI                \\ \hline
\textbf{Axis}                                    & 3                & 1 or 3             \\ \hline
\textbf{Range} (g)                               & $\pm$ 3          & $\pm$ 2 to 16      \\ \hline
\textbf{Bandwidth} (kHz)                         & 0.55             & 5 - 6.3            \\ \hline
\textbf{Sensitivity}                             & 300 mV/g         & 0.061 mg/LSB       \\ \hline
\textbf{Noise density} ($\mu g / \sqrt{\mathrm{Hz}}$ rms) & 150 - 300        & 75                 \\ \hline
\textbf{Microcontroller}                         & Beaglebone Black & ESP32-PoE-ISO      \\ \hline
\textbf{CPU SoC}                                 & TI Sitara AM3358 & ESP32-WROOM-32     \\ \hline
\textbf{Output data rate} (kHz)                  & 2.5              & 26.7               \\ \hline
\textbf{ADC resolution} (bit)                    & 12               & 16                 \\ \hline
\textbf{FIFO}                                    & -                & 3 kB (512 samples) \\ \hline
\end{tabular}
\caption{Hardware parameters of previous and proposed sensor units}
\label{tab:design:hw-sensors}
\end{table}

The IIS3DWB accelerometer is a cheap enough MEMS accelerometer that approaches industrial standards for vibration monitoring. An interface capable of higher speeds is needed to store all acceleration samples. Therefore, ESP32-PoE-ISO is chosen as a microcontroller development kit because it has an SD card slot connected via SD/MMC bus. The accelerometer uses an SPI bus with a maximum speed of 10 MHz that can be connected to any physical GPIO pins. The block diagram of the designed hardware device is in Figure~\ref{fig:design:block-diagram-hw}.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.9\textwidth]{assets/design/hw-block-schematic.png}
	\caption{Sensor unit hardware block diagram}
	\label{fig:design:block-diagram-hw}
\end{figure}

The drivers for necessary low-level interfaces and FAT32 filesystem are  implemented already in ESP-IDF SDK\footnote{ESP-IDF SDK: \url{https://docs.espressif.com/projects/esp-idf/en/latest/esp32/api-reference/index.html}}. The accelerometer driver is also made available by the vendor\footnote{IIS3DWB driver: \url{https://github.com/STMicroelectronics/iis3dwb-pid}}.

\section{Preliminary measurements}
During the initial on-site inspections of machinery, we carried out preliminary measurements of vibration signals. Raw waveforms of acceleration for the fan, two different compressors, and outer bearing on a pump in three axes are plotted in Figure~\ref{fig:design:preliminary-acceleration}. The acceleration rarely exceeds 1~g.

In frequency spectra in Figure~\ref{fig:design:preliminary-spectrum} behavior of machines can be clearly discerned, below the resonance frequency of the accelerometer (550 Hz). The compressor suspected to be faulty has stronger \nth{4} harmonic and above than the other one.

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{assets/design/EDA-custom-dataset-temporal.png}
        \caption{Acceleration in $\mathrm{m/s}^2$}
        \label{fig:design:preliminary-acceleration}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{assets/design/EDA-custom-dataset-velocity.png}
        \caption{Approximate velocity in $\mathrm{mm/s}$}
        \label{fig:design:preliminary-velocity}
    \end{subfigure} 
    \caption{Temporal domain waveforms in all spatial directions with 300 ms duration at 1 s timestamp}
\end{figure}

Figure~\ref{fig:design:preliminary-velocity} documents the effort to obtain vibration velocity (as per ISO~20186) by integration with the composite trapezoidal rule. The significant drift after integration is eliminated by subtracting the mean of the envelope constructed using cubic interpolation of peaks. Envelopes must be tweaked further to achieve more accurate values.

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{assets/design/compressor-orbitals-1x.png}
		\caption{Orbital analysis of compressor's \nth{1} harmonic frequency}
		\label{fig:design:orbital-analysis}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.54\textwidth}
        \includegraphics[width=\textwidth]{assets/design/EDA-custom-dataset-spectral-X-axis.png}
        \caption{FFT with Hann window of length $2^{10}$ ($\approx 409\;\mathrm{ms}$)}
        \label{fig:design:preliminary-spectrum}
    \end{subfigure} 
    \caption{Orbital and spectral methods of vibration signal analysis}
\end{figure}

The orbital plot of the first harmonic frequency $f_{max} \pm 5\;\mathrm{Hz}$ in Figure~\ref{fig:design:orbital-analysis} allows us to check the sensor deviation out of the axis of rotation.

### 2023-12-27 thesis/chapters/preliminaries/annotation.tex
Move machinery and sensors to implementation and resume background

Z vibračných signálov sú odvodené trendové indikátory a tak vytvárajú sumarizujúce číselné atribúty. Podobnosť atribútov so závislou premennou popisujúcou typy porúch sa posudzuje metrikami výberu atribútov v učení s učiteľom. Metriky sú hodnotené individuálne a súborovo. Presnosť detekcie porúch s algoritmom k-najbližších susedov sa vyhodnocuje na podskupinách najlepších atribútov, aby sa znížili požiadavky na objem prenosu dát v monitorovacom riešení. Model prírastkového učenia sa porovnáva s jeho dávkovým náprotivkom, v situáciach kedy je expertné označenie prevádzkového stavu oneskorené alebo chýbajúce.

Na základe súboru údajov MaFaulDa sú klasifikačné metriky vyhodnocované podľa viacerých scenárov a tiež sú vyjdrené stratové kompresné pomery. Plán merania a senzorové zariadenie sú navrhnuté na zber nového súboru údajov o správaní sa strojov ako sú kompresory a čerpadlá. Prínos tejto práce spočíva v identifikácii možnosti použitia iba pár veličín na opis nielen prítomnosti poruchy, ale aj jej príčiny.
\emptypage

>>>>

Z vibračných signálov sú odvodené trendové indikátory a tak vytvárajú sumarizujúce číselné atribúty. Podobnosť atribútov so závislou premennou popisujúcou typy porúch sa posudzuje metrikami výberu atribútov v učení s učiteľom. Metriky sú hodnotené individuálne a súborovo. Presnosť detekcie porúch algoritmom k-najbližších susedov sa vyhodnocuje na podmnožinách najlepších atribútov. Tým sa znížia požiadavky na objem prenosu dát v monitorovacom riešení. Model prírastkového učenia sa porovnáva s učením v dávkach, v situáciach kedy je expertné označenie prevádzkového stavu oneskorené alebo chýbajúce.

Na základe súboru údajov MaFaulDa sú klasifikačné metriky vyhodnotené podľa viacerých scenárov a tiež sú vyjdrené stratové kompresné pomery. Plán merania a senzorové zariadenie sú navrhnuté na zber nového súboru údajov o správaní sa strojov ako sú kompresory a čerpadlá. Prínos tejto práce spočíva v identifikácii možnosti použitia iba pár veličín na opis nielen prítomnosti poruchy, ale aj jej príčiny.
\emptypage
